---
title: "EDA LA-1"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
date: '2022-06-23'
---
### 1.Reading the data from the dataset
```{r}
data=read.csv("C:/Users/Monisha/OneDrive/Documents/Rstudio/R/COVID-19 India Statewise Vaccine Data.csv")
```

### 2.To view all objects
```{r}
ls()
```

### 3.Using ls() for pattern matching
```{r}
ls(data)
```

### 4.Finding the class of dataset
```{r}
class(data)
```

### 5.Finding the structure of dataset
```{r}
str(data)
```

### 6.Selectind the element present in 3rd row and 3rd column
```{r}
data[3,3]
```

### 7.Selecting 3rd row and displaying columns 1 to 4
```{r}
data[3,1:4]
```

### 8.Selecting the first column alone
```{r}
data[,1]
```

### 9.Displaying specific rows and all columns
```{r}
data[c(1,3),]
```
### 10.Displaying specific rows and all columns except 4th column
```{r}
data[c(1,3),-4]
```
### 11.Fetching column data by its name
```{r}
data[c(1,3),"State.UTs"]
```
### 12.Sorting the column 
```{r}
sort(data$State.UTs)
```
### 13.Ordering a particular column in a dataset
```{r}
order(data$Population)
```
### 14.Determine the rank of particular column
```{r}
rank(data$Dose.1)
```
### 15.Displaying first six rows of the dataset
```{r}
head(data)
```
### 16.Displaying the last six rows of the dataset
```{r}
tail(data)
```
###17.Displaying first 3 rows in a dataset
```{r}
head(data,n=3)
```
### 18.Summary of dataset
```{r}
summary(data)
```
### 19.To display column names
```{r}
colnames(data)
```
### 20.To display row names
```{r}
row.names(data)
```
### 21.To display both row and column names
```{r}
dimnames(data)
```
### 22.To find the transpose of the data frame
```{r}
t(data)
```
### 23.Ordering the rows based on column name
```{r}
with(data,order(State.UTs))
```
### 24.Converting dataframe to a matrix
```{r}
as.matrix(data)
```
### 25.Converting dataframe to list
```{r}
as.list(data)
```
### 26.To get the 20% quantile of a given column
```{r}
quantile(data$Population,0.2)
```
### 27.To pick 20%,50%, and 80% quantile
```{r}
quantile(data$Population,c(0.2,0.5,0.8))
```
### 28.Trying quantiles for non numeric order data
```{r}
quantile(data$Population, c(0.5,0.75,0.25))
```
### 29.To determine the cumulative sum of data
```{r}
cumsum(data$Dose.1)
```
### 30.To get the cumulative maximum value 
```{r}
cummax(data$Dose.1)
```
### 31.To get the minimum cumulative sum
```{r}
cummin(data$Dose.1)
```
### 32.To get the cumulative product of the sample
```{r}
cumprod(data$Dose.2)
```
### 33.Trying cumulative command on a vector of character data
```{r}
cummax("State.UTs")
```
### 34.To generate sequences of values
```{r}
seq(data$Dose.2)
```
### 35.By using along we are generating a vector
```{r}
seq(along=data$Dose.2)
```
### 36.To get the mean of column 2
```{r}
mean(data[,2])
```
### 37.To get the sum of each column
```{r}
colSums(data[,c(2,3)])
```
### 38.To get the mean of each row
```{r}
rowMeans(data[,c(3,4)])
```
### 39. Apply command
```{r}
apply(data[,c(3,4)],2,median)
```

### 40.Max of a column
```{r}
max(data$Population)
```
### 41.lapply which is used to apply for all columns
```{r}
lapply(data,mean,na.rm=TRUE)
```
### 42.sapply same as lapply in a prettier form
```{r}
sapply(data,mean,na.rm=TRUE)
```
### 43.Creating list of column names
```{r}
colnames=names(data)
colnames
```
### 44.Creating contingency table of dataframe
```{r}
data.tab = with(data,table(colnames[1],colnames[2]))
data.tab
```
### 45.Display the table as proportions of the row tables
```{r}
prop.table(data$Dose.2)
```

### 46.Create a stem and leaf plot for the dataframe
```{r}
stem(data$Population)
```
### 47.Loading dplyr package to use predefined functions to perform on dataset
```{r}
library(dplyr)
```
### 48.Adding a column to the dataset
```{r}
data %>% mutate(newcol = NA)
```
### 49.Renaming the first column as states
```{r}
data = data %>% rename(states = "State.UTs")
data
```
### 50.Getting maximum value in Dose.1 column
```{r}
max(data$Dose.1)
```
### 51.Reordering columns in a dataset
```{r}
data %>% select(Dose.2,Dose.1,Population)
```
### 52.Getting a subset of a dataframe
```{r}
data[data$bookft == "Population" &
 data$reviews >= 1 &
 data$reviews <= 20, 
 c("Dose.1", "Dose.2")]
```
### 53.Loading ggplot2 library
```{r}
library(ggplot2)
```
### 54.To make a scatter plot where x is states and y is Dose.1
```{r}
ggplot(data,aes(x=states,y=Dose.1))+geom_point()
```

### 55.To make line graph
```{r}
plot(data$Dose.1,data$Population,type="l")
```

### 56.Plotting line graphs along with points
```{r}
plot(data$Dose.1,data$Population,type="l")
points(data$Dose.1,data$Population)
```

### 57.Comparing dose1 and dose2 data by plotting line point graph
```{r}
plot(data$Dose.1,data$Population,type="l")
points(data$Dose.1,data$Population)
plot(data$Dose.2,data$Population,type="l",col="red")
points(data$Dose.2,data$Population,col="red")
```
### 58.Comparing geom_line and geom_point
```{r}
ggplot(data,aes(x=states,y=Dose.1))+geom_line()
ggplot(data,aes(x=states,y=Dose.1))+geom_line()+geom_point()
```
### 59.Bar graph of data frame
```{r}
ggplot(data,aes(x=states,y=Population))+geom_col()
```
### 60.Converting x coordinate into factor and then plotting bar graph so that it is discrete
```{r}
ggplot(data,aes(x=factor(Dose.1),y=Dose.2))+geom_col()
```
### 61.To plot bar graph
```{r}
ggplot(data,aes(x=Dose.1))+geom_bar()
```
### 62.To plot bar graph using factor
```{r}
ggplot(data,aes(x=factor(Dose.1)))+geom_bar()
```
### 63.Plotting Histogram
```{r}
hist(data$Dose.1)
```
### 64.Plotting histogram of dose2
```{r}
hist(data$Dose.2)
```
### 65.Plotting histogram of population
```{r}
hist(data$Population)
```
### 66.Plotting histogram with specifying approximate number of bins with breaks
```{r}
hist(data$Dose.1,breaks=10)
```
### 67.Using ggplot to plot histogram
```{r}
ggplot(data,aes(x=Dose.1))+geom_histogram()
```

### 69.Creating boxplot
```{r}
boxplot(data$Population,data$Dose.1)
```
### 70.Plotting for population and Dose1
```{r}
plot(data$Population,data$Dose.1)
```
### 71.Plotting box plot with 2 variables in x-axis
```{r}
boxplot(Dose.1~Dose.2,data=data)
```
### 72.Plotting boxplot by putting interaction of two variables on x-axis
```{r}
boxplot(Population~Dose.1+Dose.2,data=data)
```
### 73.Plotting boxplot using ggplot2 package
```{r}
ggplot(data,aes(x=states,y=Population))+geom_boxplot()
```
### 74.Plotting boxplot using ggplot for multiple variables
```{r}
ggplot(data,aes(x=interaction(Dose.1,Dose.2),y=Population,states))+geom_boxplot()
```
## 75.Plotting a boxplot using color fill
```{r}
ggplot(data,aes(x=Dose.1,y=Dose.2))+geom_col(fill="lightblue",colour="red")
```
### 76.Using mapcultivator to fill color into bargraph
```{r}
ggplot(data,aes(x=Dose.1,y=Dose.2,fill = states))+geom_col(position="dodge")
```
### 77.Using mapcultivator to fill color into bargraph for population
```{r}
ggplot(data,aes(x=Dose.1,y=Dose.2,fill = Population))+geom_col(position="dodge")
```
### 78.Usin scale_fill_brewer() to the boxplot for changing colors
```{r}
ggplot(data,aes(x=Dose.1,y=Dose.2,fill=states))+geom_col(position="dodge",colour = "green")+scale_fill_brewer(palette = "Pastel1")
```
### 79.Rearranging and slicing 
```{r}
ans = data %>%
  arrange(desc(data)) %>%
  slice(1:10)
ans
```
### 80.Then plotting the above data
```{r}
ggplot(ans,aes(x=states,y=Dose.1,fill=Population))+geom_col()
```
### 81.Plotting bar graph without spaces between bars
```{r}
ggplot(data,aes(x=states,y=Dose.2,fill=Population))+geom_col(width=0.5,position = "dodge")
```
### 82.Plotting bar graph with spaces between bars
```{r}
ggplot(data,aes(x=states,y=Dose.2,fill=Population))+geom_col(width=0.5,position = position_dodge(0.7))
```
### 83Reversing the stacking order of the graph
```{r}
ggplot(ans,aes(x=states,y=Dose.1,fill=Population))+geom_col()+guides(fill=guide_legend(reverse=TRUE))
```
### 84.Adding text to x and y axis
```{r}
ggplot(ans, aes(x = interaction(Dose.1, Dose.2), y = Population)) +
 geom_col() +
 geom_text(aes(label = Population), vjust = 1.5, colour = "green")
```
### 85.Adding text to x and y axis above the bars
```{r}
ggplot(ans, aes(x = interaction(Dose.1, Dose.2), y = Population)) +
 geom_col() +
 geom_text(aes(label = Population), vjust = -0.2, colour = "green")
```
### 86.Putting the label in the middle of each bar
```{r}
ce <- ans%>%
 group_by(Dose.1) %>%
 mutate(label_y = cumsum(Dose.1) - 0.5 * Dose.1)
ggplot(ce, aes(x = states, y = Dose.1, fill = Dose.2)) +
 geom_col() +
 geom_text(aes(y = label_y, label = Dose.1), colour = "green")
```
### 87.Setting the dose 2 column in the range from 0 to max
```{r}
ggplot(ans, aes(x = Dose.1, y = Dose.2)) +
 geom_line() +
 ylim(0, max(ans$Dose.2))
```
### 88.Setting the dose 2 column in the range from 0 to max by using expand_limits()

```{r}
ggplot(ans, aes(x = Dose.1, y = Dose.2)) +
 geom_line() +
 expand_limits(y = 0)
```
### 89.Plotting line and point graph for ans data

```{r}
ggplot(ans, aes(x = Dose.1, y = Dose.2)) +
 geom_line() +
 geom_point()
```
### 90.Plotting the line and point graph with inconsistent data points
```{r}
ggplot(ans, aes(x = Dose.1, y = Population)) +
 geom_line() +
 geom_point()
```
### 91.Plotting the line and point graph with inconsistent data points and log on y axis
```{r}
ggplot(ans, aes(x = Dose.2, y = Population)) +
 geom_line() +
 geom_point() +
 scale_y_log10()
```
### 92.Adding mapsup colour to line graph
```{r}
ggplot(ans, aes(x = Dose.2, y = Dose.1, colour = Population)) +
 geom_line()
```
### 93.Mapping variables to propertoes of a point
```{r}
ggplot(ans, aes(x = Dose.1, y = Dose.2, shape = states)) +
 geom_line() +
 geom_point(size = 4)
```
### 94.Mapping variables to propertoes of a point of different colors and same shape
```{r}
ggplot(ans, aes(x = Dose.1, y = Dose.2, fill = states)) +
 geom_line() +
 geom_point(size = 4, shape = 21)
```

### 95.Plotting dashed line graph
```{r}
ggplot(ans, aes(x = Dose.1, y = Dose.2)) +
 geom_line(linetype = "dashed", size = 1, colour = "blue")
```

### 96.Plotting dashed lined raph with points
```{r}
ggplot(ans, aes(x = Dose.1, y = Dose.2, colour = states)) +
 geom_line(linetype = "dashed") +
 geom_point(shape = 22, size = 3, fill = "white")
```

### 97.Plotting areaplot
```{r}
ggplot(ans, aes(x = Population, y = Dose.1)) +
 geom_area()
```

### 98.Plotting areaplot with colours for better analysis
```{r}
ggplot(ans, aes(x = Dose.1, y = Dose.2)) +
 geom_area(colour = "black", fill = "blue", alpha = .2)
```

### 99.Area graph With a dotted line for upper and lower bounds
```{r}
ggplot(ans, aes(x = Dose.1, y = Dose.2)) +
geom_line(aes(y = Dose.1 - Dose.2), colour = "grey50",
linetype = "dotted") +
geom_line(aes(y = Dose.1 + Dose.2), colour = "grey50",
linetype = "dotted") +
geom_line()
```

### 100.Plotting line graph on top of shaded regions
```{r}
ggplot(ans, aes(x = Dose.1, y = Dose.2)) +
geom_ribbon(aes(ymin = Dose.2 - Dose.1,
ymax = Dose.1 + Dose.2),
alpha = 0.2) +
geom_line()
```

